Numerical trajectory patterns of floating macroalgae

### 2021-02-19 thesis_chapter_3.Rmd
chapter CR additions and extra additions

There are a range of objects, both natural and anthropogenic, floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa. Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]. Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species of macroalgae have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant. In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum. The structure and number of pneumatocysts varies between species. For example species from the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018]. In some cases the stipe itself is air-filled, such as with *E. maxima* and *E. radiata*. Although there have been reports of floating Chlorophyta species and some Rhodophyta species, Phaeophyceae species are the most commonly reported forms of floating algae. This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped inbetween or in the thalli [@dromgoole1982; @back2000]. The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018] and the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.

Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognised as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterrey Bay. The results showed that kelp rafts, with little surface, area exposed to the wind were largely driven by a combination of wind and wind waves, however the relative importance of wind and wind waves was not clear. In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter. Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example, the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For example, a study by @fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft. The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate; while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011].

Other aspects such as buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material. However, the past research on macroalgal trajectory has not investigated these factors which have been shown to be important aspects of trajectory for other materials such as icebergs, marine craft and microplastics. The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag. Drag coefficients related to wind and surface current have been shown to be important properties to consider when estimating trajectory and forecasting drift for search and sea rescue operations. Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species. In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object. If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].

@@

The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A). A release site was chosen for this study, Kommetjie, which is located on the Atlantic side of the Cape Peninsula (Figure 2B). The hydrodynamics in the study region are driven by a complex interaction of wind and waves, which are also influenced by larger oceanic processes. In South Africa, the larger processes can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC)[@lutjeharms2000;@lutjeharms2007;@garzoli1996;@Hutchings2009;@rubio2009]. The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@@lutjeharms2000;@lutjeharms2007]. The region where the retroflection occurs is known as the Agulhas retroflection and consists of some of the highest levels of mesoscale variability [@garzoli1996]. The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @lutjeharms2007]. The Agulhas Rings are the main contributors to the Agulhas leakage which are the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015]. The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009]. The Benguela region is characterised by strong coastal upwelling cells which are primarily driven by equatorward winds [@shannon1996; @hardman2003], and is known as the Benguela upwelling region. The Benguela upwelling region starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996]. The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments [@blanke2002; @blanke2005; @rubio2009]. The Benguela inshore region is dominated by a northwestward flow as a result of topographical steering, wind stress and interactions with passing Agulhas rings and eddies [@veitch2010].

```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1:Map of the study domain. This is a temporary map until issues with the map script are resolved."}
library(png)
library(grid)
library(gridExtra)

#map_chp3A <- rasterGrob(as.raster(readPNG("map/chp3_mapA.png")))
#map_chp3B <- rasterGrob(as.raster(readPNG("map/chp3_mapB.png")))
#ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_combined_hz.png")))
ggarrange(map_chp3_combined)
```

@@

Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'. All simulations have the same release location and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018]. Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model. These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018]. To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios. These scenarios were meant to act as proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag. These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation. For example, if 85% of the plant is submerged and 15% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively. The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1).

```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Details for each simulation run for the study"}
set_flextable_defaults(fonts_ignore=TRUE)

@@

sim_table <- flextable(data = sim_table)
sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
sim_table <- align(x= sim_table, align = "center", part = "all")
sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
sim_table <- set_caption(sim_table, "Table 1:Details for each simulation executed for the study.")
sim_table
```

@@

The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical. The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field. Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors. Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step. The same approach was used for the wind drag force. The equation that was used to calculate hydrodynamic drag, was also used to calculate wind.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6, fig.cap="Figure 2: Comparison of mean trajectories. Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag. In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10. In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13."}
ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
```

@@

The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020]. Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], current research has focused on large macroalgae-rafts which can vary greatly in size and surface area. None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae. This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag. The mean virtual kelp trajectory deviated from the mean passive  trajectory when including varying levels of hydrodynamic and wind drag scenarios. The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns, and also our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.

@@

The addition of drag forces into the simulation also causes the particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed. This is most likely due to the shared drag characteristics among the particles which included the plants weight and surface area into the calculation. Other studies investigating flotsam trajectory characteristics have found similar results [@breivik2011; @olascoaga2020; @miron2020]. A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, cuboids to compare to investigate the effect of inertia on particles dynamics. The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia. Although drag forces were used in this study and not inertia, the same conclusion applies, drag is an important characteristic to consider when investigating or predicting solitary drifting macroalgae.

The ocean is made up of varying time-varying flows and changes in an objects velocity can cause the object to be exposed to different flow patterns over time. The lack of dissimilarity between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible. A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region. The results suggested that the hatchlings trajectories are mostly influenced by the ocean currents in the first year of hatching due to low swimming speeds of juveniles. Therefore, the differences in cross-sectional area types may not be significant enough in order to have a significant effect on the relevant velocity vectors (zonal and meridonal). Another possible reason is the resolution of the ocean model used. The resolution of the underlying ocean model is an important aspect of lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications. A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study. The authors found that the inclusion of brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications. Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.

The same is also true when comparing different hydrodynamic and wind drag exposure scenarios. These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag. The similarities between the different combined drag scenarios suggests that higher magnitudes of wind exposure do not significantly alter trajectories. Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns. All the simulations that included any form of drag resulted in a fairly high density of particles entrained in a vortex. The entrainment and expulsion from eddies and vortices is an important characteristic of floating macroalgae. Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts. However, the aforementioned studies investigated large rafts and not floating individuals. Inertia is based on the size/weight of the raft, while hydrodynamic and wind drag are based on current velocity and surface area. Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag should be considered. The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices. The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory. Instead, the inclusion of a different shape coefficient increased to the density of particles entrained in an eddy field. This suggest that inclusion of different shape coefficients play a role in further improving estimates and hindcasts of macroalgal dispersal patterns.

>>>>

There are a range of objects, both natural and anthropogenic, floating in the ocean of which macroalgae are regarded as one of the most important passive dispersal mechanisms of marine taxa. Floating kelp acts as a passive dispersal mechanism for a range of marine taxa and is sometimes referred to as the 'tumble-weed' of the ocean [@edgar1987; @bushing1994; @norton1992; @helmuth1994; @holmquist1994; @smith2002; @McCormick2008]. Some macroalgae species are negatively buoyant and sink to the seafloor when detached from the substrate, while other species of macroalgae have air-filled pneumatocysts or stipes which allow the plants to reach the surface where light is more abundant. In turn, the positively buoyant pneumatocysts cause plants to float to the surface when dislodged from the substratum. The structure and number of pneumatocysts varies between species. For example species from the genera *Macrocystis*, *Sargassum*, *Ascophyllum*, and *Fucus* have thalli with many small pneumatocysts, while other kelp species such as *Nereocystis luetkeana*, *Pelagophycus porra*, *Ecklonia radiata*, *Ecklonia maxima* have a single, large pneumatocysts [@dayton1985; @smith2002; @thiel2005; @graiff2016; @batista2018]. In some cases the stipe itself is air-filled, such as with *E. maxima* and *E. radiata*. Although there have been reports of floating Chlorophyta species and some Rhodophyta species, Phaeophyceae species are the most commonly reported forms of floating algae. This is most likely because the green and red species reported floating are not actually positively buoyant, but instead are kept at the surface by gas trapped inbetween or in the thalli [@dromgoole1982; @back2000]. The giant kelp *Macrocystis pyrifera* [@helmuth1994; @kingsford1995; @hobday2000; @macaya2005; @graiff2016; @batista2018], the bull kelp *Durvillaea antarctica* [@collins2010; @tala2013 ; @tala2017; @saunders2014; @smith2002; @wichmann2012; @batista2018], and pelagic *Sargassum* species [@putman2020; @brooks2019; @putman2018; @brooks2018] have been the focus of much of the research regarding spatial and temporal dispersal patterns, the ecological role of rafting, marine connectivity and raft-time.

Past research points to macroalgae trajectory being largely determined by prevailing wind conditions and surface currents [@thiel2005; @hobday2000; @fraser2011; @rothausler2011; @rothausler2011]. Although ocean currents are regarded as the primary influence, the relative importance of wind versus surface current is still unknown although the role of wind has been recognised as important in several studies. For example, a study by @harrold1989 investigated the seasonal trajectories of radio-transmitter tagged *M. pyrifera* in nearshore Monterrey Bay. The results showed that kelp rafts, with little surface, area exposed to the wind were largely driven by a combination of wind and wind waves, however the relative importance of wind and wind waves was not clear. In addition, the tagged kelp trajectories were more consistent with the formation of eddies during winter. Previous studies have identified wind as an important mechanism of dispersal in wind dominated ocean systems. For example, the subAntartic latitudes the West Wind Drift causes continuous unidirectional surface flow and is regarded as an important potential mechanism for dispersal of floating kelp. Other studies have used genetic approaches to determine macroalgae raft trajectory characteristics by inferring source location from genetically similar populations [@nikula2013]. For example, a study by @fraser2011 on the rafting capabilities of *Durvillaea antarctica* used a combination of population genetics and relative age estimate of ‘goosebarnacles’ attached to the raft. The presence of goosebarnacles suggests a long raft time as these species have a slow growth rate; while the genetic analyses showed these species are able to raft up to $\sim$ 390km from their local origin. The authors suggested that wind and water-movement were the primary influences of trajectory, however, this was only inferred from the genetic results and local climatology data [@fraser2011]. Past research has shown that buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material [onink2019; @shen2019; @dasaro2018; @gates2018; @beron2016; @trinanes2016; @furnans2008]. The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in hydrodynamic form drag [@breivik2011; @griffin2017]. In addition, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as drivers of trajectory, as well as orientation of the object. If the length of the object is longer than the significant wave height then waves will be the primary driver of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017]. Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae species.

@@

The region of interest for this study was the Cape Peninsula in the Western Cape (Figure 1A). A release site was chosen for this study, Kommetjie, which is located on the Atlantic side of the Cape Peninsula (Figure 2B). The hydrodynamics in the study region are driven by a complex interaction of wind and waves, which are also influenced by larger oceanic processes. In South Africa, the larger processes can largely be attributed to the Agulhas Current (AC) and the Benguela Current (BC)[@lutjeharms2000;@lutjeharms2007;@garzoli1996;@Hutchings2009;@rubio2009]. The AC is a component of the South-west Indian Ocean sub-gyre, flowing predominantly south-westward following the continental shelf edge and eventually retroflecting eastward back into the South Indian Ocean [@@lutjeharms2000;@lutjeharms2007]. The region where the retroflection occurs is known as the Agulhas retroflection and consists of high levels of mesoscale variability [@garzoli1996]. The variability results in shedding of rings and filaments known as the Agulhas Rings and have an average diameter of ~324km [@lutjeharms1988; @lutjeharms2006; @lutjeharms2007]. The Agulhas Rings are the main contributors to the Agulhas leakage which are the transport of warm water from the Indian Ocean to the South Atlantic Ocean [@garzoli1996; @lutjeharms2006; @beal2015]. The Agulhas leakage is also a source of meso-scale variability within the Benguela region as anticyclonic rings are shed and drift northwestward along the southern Benguela slope region [@rubio2009]. The Benguela region is characterised by strong coastal upwelling cells which are primarily driven by equatorward winds [@shannon1996; @hardman2003], and is known as the Benguela upwelling region. The Benguela upwelling region starts at 27$^\circ$S and extends southward to 35$^\circ$S [@shannon1996]. The region is also characterised by high amounts of mesoscale variability in the form of eddies and filaments [@blanke2002; @blanke2005; @rubio2009]. The Benguela inshore region is dominated by a northwestward flow as a result of topographical steering, wind stress and interactions with passing Agulhas rings and eddies [@veitch2010].

```{r map, echo=FALSE, message=FALSE, warning=FALSE, strip.white=TRUE, fig.cap="Figure 1:Map of the study domain. This is a temporary map until issues with the map script are resolved."}

library(png)
library(grid)
library(gridExtra)

#map_chp3A <- rasterGrob(as.raster(readPNG("map/chp3_mapA.png")))
#map_chp3B <- rasterGrob(as.raster(readPNG("map/chp3_mapB.png")))
#ggarrange(map_chp3A, map_chp3B, ncol = 2, nrow = 1, labels = "AUTO", align = c("h"))
map_chp3_combined <- rasterGrob(as.raster(readPNG("map/chp3_map_combined_hz.png")))
ggarrange(map_chp3_combined)

The study domain falls within the southern Benguela upwelling ecosystem, which is a region of high mesoscale variability and activity. These occur mostly in the form of eddies, filaments and vortices which are generated from the Agulhas retroflection [@veitch2018; @veitch2017];  generally flowing northward along the coastline and begin to move offshore near Cape Columbine. The Benguela Jet is a coastal flowing current and is driven by a complex interaction of eddies, upwelling, eckman transport and geostrophic currents [@stramma1989, @veitch2017; @veitch2018; @hutchings2009]. The stability of the jet varies seasonally with more stable flows in winter and splitting of the jet during summer and spring. In January, the jet begins to split between Cape Point and Cape Columbine [@veitch2018; @blanke2009] which results into a weak narrow inshore branch which flows along the shelf edge and a strong flowing north-westwards offshore branch. The area is also an area of high upwelling activity driven by strong westerly winds [@hutchings2009]. These winds are most intense during summer (December-March) which favours upwelling conditions in the region, and also act to reinforce the flow of the Benguela Jet; which occurs due to cold upwelling waters contributing to the across-shore density gradient [@ragoasha2019]. In winter the wind direction is reversed which decreases upwelling favorable conditions, but the core of the Benguela jet is maintained by the intrusion of warm agulhas waters into the Benguela region over the shelf edge [@veitch2018; @hutchings2009; @ragoasha2019].

@@

Virtual particles were released with no drag or wind effects from the study site in order to act as a reference for comparison to simulations with varying degrees of hydrodynamic and wind drag, and are referred to hereafter as 'passive particles'. All simulations have the same release location and were run for 30 days with hourly outputs using a fourth order Runge-Kutte advection scheme [@North2009; @VanSebille2018]. Brownian motion was also incorporated in order to account for any stochasticity not resolved by the model. These are standard practices for Lagrangian trajectory studies [@North2009;@delandmeter2019; @VanSebille2018]. To test the different magnitudes that hydrodynamic and wind drag might play in the overall trajectory, simulations were run with varying degrees of drag exposure scenarios. These scenarios were meant to act as proxy for buoyancy, which determines the surface area exposed to hydrodynamic and wind drag. These estimates were expressed as percentages which were used to calculate hydrodynamic and wind drag for applicable simulation. For example, if 85% of the plant is submerged and 15% of above water, then the overall hydrodynamic and wind drag forces would be 85% and 15% of the total drag forces respectively. This approach has been used in past research investigating pelagic Sargassum rafts [@putman2020;@putman2018]. The exposure scenarios used in this study were 100%, 95%, 90%, and 85% for hydrodynamic drag; 0 %, 5%, 10%, and 15% for wind (see table 1). Each scenario consisted of a particle set with a size of 1000 where released from Kommetjie at the start of simulation. The particles sets were advected with a Runge-Kutte fourth order scheme for 30 days and incorporated brownian motion. Since the aim of this investigation was to investigate the relative surface behaviour of virtual kelp particles, only the surface layer was used for simulations. It should also be noted that the Copernicus model used places velocity vectors over some of the land covering the coastal domain. This was not considered a concern, as once again, the aim was to investigate surface drift behavior. The time period chosen to run the simulations was 01-01-2018 un 30-01-2018. This particular time period was chosen as the splitting of the Benguela Jet occurs in Janurary, which also conincides with a period of higher wind intensity, upwelling along the coast and mesoscale activity [@rubio2009; @veitch2018; @ragoasha2019].

```{r sim table, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Details for each simulation run for the study"}

@@

sim_table <- flextable(data = sim_table)
sim_table <- theme_zebra(sim_table, odd_header = "transparent", even_header = "transparent")
sim_table <- set_header_labels(sim_table, Cross.sectional.area = "Cross-sectional area", Drag.Exposure = "Drag Exposure level")
sim_table <- align(x= sim_table, align = "center", part = "all")
sim_table <- set_table_properties(sim_table, width = 1, layout = "autofit")
sim_table <- set_caption(sim_table, "Table 1:Details for each simulation executed for the study.")
sim_table

@@

The drag coefficient for a spherical particle was used in the calculation of hydrodynamic drag as the trajectory model assumes the particles are spherical. The drag equation used was: $F_d = 0.5 \rho A_c C_d U^2$, where $\rho$ is the density of seawater or air, $A_c$ is the cross-sectional area exposed to hydrodynamic or wind drag, $C_d$ is the drag coefficient for a spherical particle and $U$ is the surface velocity vector of the flow-/wind- field. Hydrodynamic and wind drag components were calculated separately for both the meridonal and zonal velocity vectors, as per @furnans2008. Since drag force is dependent on the velocity vectors which vary with time, the meridonal and zonal velocities were interpolated and used in the drag force calculation for each time-step. The same approach was used for the wind drag force. The equation that was used to calculate hydrodynamic drag, was also used to calculate wind drag.

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Comparison of mean trajectories. Plot A is the comparison of mean trajectories for only hydrodynamic drag and plot B is the comparison of mean trajectories for simulations with varying degrees of hydrodynamic and wind drag. In plot A, the black line represents simulation 1, the blue line 2, the orange line 6 and the red line 10. In plot B, the black line represents simulation 1, the grey line simulation 3, the green 4, the pink 5, the yellow 7, orange 8, red 9, blue 11, bronze 12, and cherry 13."}
ggarrange(waterdrag_plot, wwdrag_plot, labels = "AUTO")
```

@@

The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices [@fossette2012; @miron2020]. Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics [@putman2018; @brooks2018;@brooks2019; @putman2020], past research has focused on large macroalgae-rafts which can vary greatly in size and surface area. None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae. This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag. The mean virtual kelp trajectory deviated from the mean passive  trajectory when including varying levels of hydrodynamic and wind drag scenarios. The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns, and also our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.

@@

The addition of drag forces causes the particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed. This is most likely due to the shared drag characteristics among the particles which included the plants weight and surface area into the calculation. Other studies investigating flotsam trajectory characteristics have found similar results [@breivik2011; @olascoaga2020; @miron2020]. A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, cuboids to compare to investigate the effect of inertia on particles dynamics. The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia. Although drag forces were used in this study and not inertia, drag is an important characteristic to consider when investigating or predicting solitary drifting macroalgae.

The ocean is made up of varying time-varying flows and changes in an objects velocity can cause the object to be exposed to different flow patterns over time. The lack of dissimilarity between different cross-sectional area types (minimum, mean and maximum) suggests that reduction in velocity in relation to cross-sectional area is negligible. A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region. The results suggested that the hatchlings trajectories are mostly influenced by the ocean currents in the first year of hatching due to low swimming speeds of juveniles. Therefore, the differences in cross-sectional area types may not be significant enough in order to have an effect on the relevant velocity vectors (zonal and meridonal). Another possible reason is the resolution of the ocean model used. The resolution of the underlying ocean model is an important aspect of lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications. A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study. The authors found that the inclusion of brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications. Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.

The same is also true when comparing different hydrodynamic and wind drag exposure scenarios. These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag. The similarities between the different combined drag scenarios suggests that higher magnitudes of wind exposure do not significantly alter trajectories. Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns. All the simulations that included any form of drag resulted in a fairly high density of particles entrained in a vortex. The entrainment and expulsion from eddies and vortices is an important characteristic of floating macroalgae. Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real Sargassum rafts. However, the aforementioned studies investigated large rafts and not floating individuals. Inertia is based on the size/weight of the raft, while hydrodynamic and wind drag are based on current velocity and surface area. Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag, such as in this study, should be considered. The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices. The use of a cylindrical shape coefficient instead of a sphere in the calculation of drag force also had no effect on the mean trajectory. Instead, the inclusion of a different shape coefficient increased to the density of particles entrained in an eddy field. This suggest that inclusion of different shape coefficients play a role in further improving estimates and hindcasts of macroalgal dispersal patterns.
